Statically Typed String Sanitation Inside a Python

### 2014-08-13 sanitation-psp14/sanitation-psp14-short.tex
finished intro.

To avoid this problem, the program must \emph{sanitize} user input. 
For example, in this case, the developer (or, more often, a framework) might define a function \verb|sanitize| that escapes any double quotes, which are necessary in order to force uesr input to be interpreted as anything other than a string literal.
%prepends double quotes and existing backslashes  with a backslash, which SQL treats safely. Note that this function is not idempotent, so it should only be called once. Guaranteeing that user input has already been sanitized before it is used to construct a command is challenging.

@@

TODO: need to update remainder of intro ... To demonstrate this approach, we present a simply typed lambda calculus with \emph{constrained strings}; that is, a set of string types parameterized by regular expressions. If $s : \stringin{r}$, then $s$ is a string matching the language $r$. Additionally, we include an operation $\rreplace{r}{s_1}{s_2}$ which corresponds to the replace mechanism available in most regular expression libraries; that is, any substring of $s_1$ matching $r$ is replaced with $s_2$. The type of this expression is the computed, and is likely ``smaller" or more constrained than the type of $s_1$. Libraries, frameworks or functions which construct and execute commands containing input can specify a safe subset $\stringin{r_{\sf spec}}$ of strings, and input sanitation algorithms can construct such a string using ${\sf rreplace}$ or, optionally, by coercion (in which case a runtime check is inserted). We also show how this system is translated into a host language containing a regular expression library such that the safety guarantee of the extended language is preserved.

Summarily, we present a simple type system extension which ensures the absence of input sanitation vulnerabilities by statically checking input sanitation algorithms which use an underlying regular expression library. This approach is \emph{composable} in the sense that it is a conservative extension. This approach is also \emph{complementary} to existing input sanitation techniques which use string replacement for input sanitation.

@@

We are also unaware of any extensible programming languages which emphasize applications to security concerns.

@@

In this section, we present an example atlang program using our extension together
with our implementation of the extension in Atlang modulo ${\sf lreplace}$.
Note that in addition to Figure \ref{fig:regex}, regular expressions in atlang
also support the {\sf \^} character, which is complementation. Complementation
is definable in terms of Figure \ref{fig:regex} whenever $\Sigma$ is
fixed.

>>>>

To avoid this problem, the program must \emph{sanitize} user input. 
For example, in this case, the developer (or, more often, a framework) might define a function \verb|sanitize| that 
%escapes any double quotes, which are necessary in order to force uesr input to be interpreted as anything other than a string literal.
prepends double quotes and existing backslashes  with a backslash, which SQL treats safely. Note that this function is not idempotent, so it should only be called once. Guaranteeing that user input has already been sanitized before it is used to construct a command is challenging.

@@

Our type system for regular string types, $\lambdas$, is defined as an exntenstion to a simply-typed lambda calculus.
Opeartions on strings are given types according to corresponding, type-level
computations on regular expressions. A concatenation operation is defined in 
order to demosntrate this principle, but the premier operation of our system
is \emph{replacement}, which captures a rich set of input sanitation techniques
including escaped characters and encodings. In addition to defining this system,
we present a correctness-preserving compilation of $\lambdas$ into an underlying
language with ordinary strings as a regular expression library.

In addition to this theoretical contribution, we also explain how our extension
may be implemented in the Ace Programming Language, which is itself an extension
of Python -- a popular web programming language.

Summarily, we present a simple type system extension which ensures the absence
of input sanitation vulnerabilities by statically checking the correctness and 
correct use of input sanitation algorithms. This approach is \emph{composable}
in the sense that it is a conservative extension. This approach is \emph{complementary}
to exsiting input sanitation techniques that use regular expressions for input santiation.

@@

The Wyvern programming language provides a general framework for composing lanugage extensions \cite{tsl}\cite{wyvern}.
Our work identifies one particular extension, and is therefore complementary to Wyvern and related
work on extensible programming languages.
We are also unaware of any extensible programming languages which emphasize applications to security concerns.

@@

In this section, we present an example atlang program using our extension together
with our implementation of the extension in Atlang modulo ${\sf lreplace}$.
Note that in addition to Figure \ref{fig:regex}, regular expressions in atlang
also support complementation using the carot.
